---
title: "rrp"
author: "Jiaming Li"
date: "2022-11-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data processing
```{r}
if (!file.exists("activity.csv") )
    {
     dlurl <- 'http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
     download.file(dlurl,destfile='repdata%2Fdata%2Factivity.zip',mode='wb')  
     unzip('repdata%2Fdata%2Factivity.zip')
    }

# Read data
data <- read.csv("activity.csv")  
```


## Mean total number of steps taken per day
```{r}
steps_by_day <- aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="blue",xlab="Number of Steps")
```

```{r}
rmean <- mean(steps_by_day$steps)
rmean
```

```{r}
rmedian <- median(steps_by_day$steps)
rmedian
```

## Average daily activity pattern

```{r}
steps_by_interval <- aggregate(steps ~ interval, data, mean)
plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```


```{r}
max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
max_interval
```


## Imputing missuing values

Total number of missing values in the dataset
```{r}
NATotal <- sum(!complete.cases(data))
NATotal
```

Using Mean for the day compute missing values
```{r}
StepsAverage <- aggregate(steps ~ interval, data = data, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(data)) {
    obs <- data[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepsAverage, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}
```

Create a new dataset including the imputed missing values
```{r}
new_activity <- data
new_activity$steps <- fillNA
```


Show the result
```{r}
StepsTotalUnion <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)
hist(StepsTotalUnion$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
#Create Histogram to show difference. 
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "green"), lwd=10)
```

Calculate mean
```{r}
rmeantotal <- mean(StepsTotalUnion$steps)
rmeantotal
```

Calculate median
```{r}
rmediantotal <- median(StepsTotalUnion$steps)
rmediantotal
```

Compare
```{r}
rmediandiff <- rmediantotal - rmedian
rmediandiff
```

```{r}
rmeandiff <- rmeantotal - rmean
rmeandiff
```

The mean(Mean Var: 0) is the same however the median does have a small variance(Median Var:1.1886792). between the total which includes the missing values to the base. \

On observation the impact of the missing data has the biggest effect on the 10000 - 150000 step interval and changes frequency from 27.5 to 35 a variance of 7.5\

## Differences in activity patterns between weekdays and weekends

Created a plot to compare and contrast number of steps between the week and weekend. There is a higher peak earlier on weekdays, and more overall activity on weekends.
```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
new_activity$dow = as.factor(ifelse(is.element(weekdays(as.Date(new_activity$date)),weekdays), "Weekday", "Weekend"))
StepsTotalUnion <- aggregate(steps ~ interval + dow, new_activity, mean)
library(lattice)
xyplot(StepsTotalUnion$steps ~ StepsTotalUnion$interval|StepsTotalUnion$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```











